\vskip-0.6in LOBBYING AND LEGISLATIVE UNCERTAINTY

### 2016-02-15 draft.tex
shorten references list

\textcolor{blue}{Recorded votes in legislatures (roll call data) are the most commonly used data source to measure politicians' spatial preferences. A well-established strategy is to use statistical techniques to represent patterns of legislative voting. The intuition behind these statistical models is that each roll call presents each legislator with a choice between a ``yea'' and a ``nay'' position. Legislators are presumed to vote for the position most similar to their own ideal policy position. As Clinton et al. (2004) note, it is possible to use Bayesian Markov chain simulation statistical methods to generate these preference estimates. This approach treats the unknown ideal points, bill parameters and legislators' utilities as random variables and condition upon the observed roll call data. The strategy is to impute values to these variables and estimate by regression the ideal points and bill parameters. The MCMC algorithm repeatedly performs these imputations and regressions and generates a large number of samples from the posterior density of the model parameters. Using this estimation procedure, we obtain summary statistics used for inference: legislators' mean and median ideal points as well as their 95$\%$ posterior confidence intervals. The latter are of particular interest to us, as we can use them to gauge the unpredictability of  legislators' voting behavior.}

\textcolor{blue}{We are interested in unpredictability vis-a-vis particular policy issues. Poole (2005) suggests using subsets of roll calls to establish legislators' ideal points separately on each issue. Preliminary work with our data indicates that a more statistically-robust method is to use all roll call votes in a legislative session or group of sessions to estimate an ideal point that is multidimension---that is, with one dimension for each interest group---allowing the partial pooling of information across roll calls. Multilevel modeling techniques, of which the standard ideal point estimation is a basic case, are perfectly suited to this task.}

\textcolor{blue}{We therefore estimate the model
\begin{equation}
  \Pr(y_{kj} = 1) = \text{logit}^{-1}\left( \sum_{g=1}^G \ga_k^{(g)}\left( \alpha_j^{(g)} - \beta_k^{(g)}\right) \right)
	\label{eq:v2}
\end{equation}
with roll call votes numbered $k=1,\ldots,K$, legislators numbered $j=1,\dots,J$, and interest groups numbered $g=1,\dots,G$. $y_{jk}=1$ if legislator $j$ votes ``Yea'' on bill $k$; $y_{kj}=0$ if she votes ``Nay'' on bill $k$. Note that interest groups take positions on multiple roll call votes and different interest groups can take positions, often on opposing sides, on each vote.}

\textcolor{blue}{This is a logistic item response (Rasch) model as adapted by Bafumi, et. al. (2005) to the legislative context. $\alpha_j^{(g)}$ is the ideal point of legislator $j$ on issue $g$ and $\beta_k^{(g)}$ is the ideal point of a legislator who is indifferent on vote $k$ for issue $g$. $\gamma_k^{(g)}$, the discrimination parameter, is a measure of the informativeness of issue $g$ for the vote on bill $k$: a large $\ga$ indicates that legislators are very likely to vote according to their preferences, whereas a small $\ga$ means that even significant margins of preference do not contribute much to the probability that legislators will vote in favor of the bill.}

\textcolor{blue}{We are experimenting with some of the more sophisticated measures of uncertainty that are possible with a rich model such as this one; for instance the group mean for $\ga$ and constructs that combine model parameters.}

@@

\textcolor{blue}{We will also explore whether there is a way to leverage the data and methodology of Lake and Millimet (2014b) that combine the contributions and lobbying data to improve precision, although there is a concern about the coarseness of the data that must be overcome.}

@@

\item Gawande, K., P. Krishna and M. Robbins (2006): ``Foreign Lobbies and U.S. Trade Policy,'' {\em Review of Economics and Statistics}, 88, 563-571.

@@

@@

	\item Buzard, K., 2014a. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.

	\item Buzard, K., 2014b. Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/2014/04/Self-enforcing-Trade_Agreements.pdf}.

	\item Conconi, P., Facchini, G., Steinhardt, M. F., Zanardi, M., 2012. The political economy of trade and migration: Evidence from the US Congress. HWWI Research Paper 136. Available at: \url{http://hdl.handle.net/10419/67944}.

	\item Dixit, A., Grossman, G., Helpman, E., 1997. Common Agency and Coordination: General Theory and Application to Government Policy Making. Journal of Political Economy 105, 752-769.

	\item Feenstra, R., 1992. How Costly is Protectionism? Journal of Economic Perspectives 6, 159-178.

	\item Gawande, K., Bandyopadhyay, U., 2000. Is Protection for Sale? Evidence on the Grossman-Helpman Theory of Endogenous Protection. The Review of Economics and Statistics 82, 139-152.

	\item Gawande, K., Krishna, P., Olarreaga, M., 2012. Lobbying Competition over Trade Policy. International Economic Review 53, 115-132.

	\item Iaryczower, M., Katz, G., Saiegh, S., 2013. Voting in the Bicameral Congress: Large Majorities as a Signal of Quality. Forthcoming in Journal of Law, Economics, and Organization.

  \item Iida, K., 1996. Involuntary Defection in Two-Level Games. Public Choice 89, 283-303.

	\item Lake, J., Millimet, D., 2014a. An Empirical Analysis of Trade-Related Redistribution and the Political Viability of Free Trade. Available at: \url{http://people.smu.edu/jlake/pdfs/VotingFTAs.pdf}

	\item Lake, J., Millimet, D., 2014b. Revisiting the link between PAC Contributions and Lobbying Expenditures. Available at: \url{http://people.smu.edu/jlake/pdfs/PACcontributionsLobbying.pdf}

	\item Milner, H., Rosendorff, B.P., 1997. Democratic Politics and International trade Negotiations: Elections and Divided Government as Constraints on Trade Liberalization. Journal of Conflict Resolution 41, 117-147.

	\item Ornelas, E., 2005. Rent Destruction and the Political Viability of Free Trade Agreements. Quarterly Journal of Economics 120, 1475-1506.

>>>>

Recorded votes in legislatures (roll call data) are the most commonly used data source to measure politicians' spatial preferences. A well-established strategy is to use statistical techniques to represent patterns of legislative voting. The intuition behind these statistical models is that each roll call presents each legislator with a choice between a ``yea'' and a ``nay'' position. Legislators are presumed to vote for the position most similar to their own ideal policy position. As Clinton et al. (2004) note, it is possible to use Bayesian Markov chain simulation statistical methods to generate these preference estimates. This approach treats the unknown ideal points, bill parameters and legislators' utilities as random variables and condition upon the observed roll call data. The strategy is to impute values to these variables and estimate by regression the ideal points and bill parameters. The MCMC algorithm repeatedly performs these imputations and regressions and generates a large number of samples from the posterior density of the model parameters. Using this estimation procedure, we obtain summary statistics used for inference: legislators' mean and median ideal points as well as their 95$\%$ posterior confidence intervals. The latter are of particular interest to us, as we can use them to gauge the unpredictability of  legislators' voting behavior.

We are interested in unpredictability vis-a-vis particular policy issues. Poole (2005) suggests using subsets of roll calls to establish legislators' ideal points separately on each issue. Preliminary work with our data indicates that a more statistically-robust method is to use all roll call votes in a legislative session or group of sessions to estimate an ideal point that is multidimensional---that is, with one dimension for each interest group---allowing the partial pooling of information across roll calls. Multilevel modeling techniques, of which the standard ideal point estimation is a basic case, are perfectly suited to this task.

We therefore estimate the model
\begin{equation}
  \Pr(y_{kj} = 1) = \text{logit}^{-1}\left( \sum_{g=1}^G \ga_k^{(g)}\left( \alpha_j^{(g)} - \beta_k^{(g)}\right) \right)
	\label{eq:v2}
\end{equation}
with roll call votes numbered $k=1,\ldots,K$, legislators numbered $j=1,\dots,J$, and interest groups numbered $g=1,\dots,G$. $y_{jk}=1$ if legislator $j$ votes ``Yea'' on bill $k$; $y_{kj}=0$ if she votes ``Nay'' on bill $k$. Note that interest groups take positions on multiple roll call votes and different interest groups can take positions, often on opposing sides, on each vote.

This is a logistic item response (Rasch) model as adapted by Bafumi, et. al. (2005) to the legislative context. $\alpha_j^{(g)}$ is the ideal point of legislator $j$ on issue $g$ and $\beta_k^{(g)}$ is the ideal point of a legislator who is indifferent on vote $k$ for issue $g$. $\gamma_k^{(g)}$, the discrimination parameter, is a measure of the informativeness of issue $g$ for the vote on bill $k$: a large $\ga$ indicates that legislators are very likely to vote according to their preferences, whereas a small $\ga$ means that even significant margins of preference do not contribute much to the probability that legislators will vote in favor of the bill.

We are currently experimenting with some of the more sophisticated measures of uncertainty that are possible with a rich model such as this one; for instance the group mean for $\ga$ and constructs that combine model parameters.

@@

\textcolor{blue}{We will also explore whether there is a way to leverage the data and methodology of Lake (2015) that combine the contributions and lobbying data to improve precision, although there is a concern about the coarseness of the data that must be overcome.}

@@

@@

\item Lake, J., 2015. Revisiting the link between PAC Contributions and Lobbying Expenditures. European Journal of Political Economy 37, 86-101.

@@
